---
title: "Data preparation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(evsim)
library(tsibble)
```


This notebook aims to describe the main objectives of this forecasting exercises and prepare the energy data sets that will be used in the next notebooks where **forecasting** will be applied. 

# Problem description

Fist of all, it is very important to describe the context and environment of our forecasting project. Answer the following questions to describe the problem appropriately:

* Who requires the forecast?​

* Which is the objective of the forecast?

* What is our dependent variable?

* What are our independent variables?

* How will we get the data?​

* Do we have access to real-time data?​

* In which frequency do we get data?​

* Is the accuracy very important?​

* Do we have to consider a specific trend in our data?

* Do we have to consider a specific seasonality in our data?

* Which predictors will have more influence over our target variable?




# Data preparation

This course aims to use custom data sets for all exercises, so below there's the code that pretends to read the data sets.

## EV sessions simulation

To simulate smart charging, the first thing we need is a data set of charging sessions. Create a custom EV model for the study case of this project using `evsim` function `evsim::get_custom_ev_model`. Remember that you have to configure the parameters of your EV user profiles in a list of tibbles (argument `parameters_lst`). You can create this list manually in this notebook or by reading the multiple workseets of an Excel file (for that you can use functions `purrr::map()`, `readxl::excel_sheets()` and `readxl::read_excel()`):

```{r}
# ev_model <- 
```


Create a custom EV charging sessions data set using `evsim` function `evsim::simulate_sessions()`. Choose the scenario you want to simulate, but a high-demand scenario is advised to better assess the impact of EV charging and the potential benefits of smart charging. Describe the corresponding chosen scenario.

- Number of sessions per day: 
- Charging power distribution: 
- Dates: 
- User profiles: 

```{r}
# ev_sessions <- 
```


Get the time-series EV demand **of every charging session** using `evsim` function `evsim::get_demand()`.

```{r}
# ev_demand <- 
```


Plot the time-series EV demand of the simulated charging sessions using `evsim` function `evsim::plot_ts()`.

```{r}

```


*Which the EV user profile with more flexibility potential? Why?* 

From now on, we will focus on this user profile to build the forecasting models and test them. Build a `tsibble` data frame with the `datetime` column as index and the `y` column with the load of the selected user profile:

```{r}
# ev_ts <- 
```


# Forecasting data

From our object `ev_ts`, the `y` column is our target or *dependent variable*. However, we have to construct our `X` matrix with all the predictors or *independent variables*. 

Mutate the `ev_ts` object with all independent variables that could influence to our objective variable:

```{r}

```



# Tidy data

For proper forecasting we should avoid the following situations:

* Gaps in data
* Including periods with behaviour that differ from the future (e.g. covid)

Check if we have any gap (i.e. `NA` values) in our data set `ev_ts`:

```{r}

```

If we do have gaps decide whether to remove the corresponding rows or fill the gaps with prevalent data. Remember that you can use functions from `tidyr` or `padr` packages.

```{r}

```


From the current variables, could you think on some transformations (e.g. log, normalization, etc.) that could help in our forecasts? 

If any, list them:

* 
* 

Now create these new features and include them in our tidy data frame.

```{r}

```



# Save data

Finally, save the environment in a file:

```{r}
save.image(file = "data.RData")
```



