---
title: "Data preparation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(purrr)
library(tidyr)
library(ggplot2)
library(feasts)
library(fable)
load("data.RData")
```


This notebook aims to perform statistic analysis to explore the data and extract insights about the most suitable type of forecasting model required in this project.

# Data exploratory analysis

## Visualization

Plot our target variable with function `ggplot2::autoplot()`, and adjust it if necessary:

```{r}

```

Use the functions from `feasts` package that you think that could provide more interesting visualizations (e.g. `gg_subseries()` or `gg_season()`). 

```{r}

```



## Statistic summary

First, print a quick summary of our data frame with function `summary()`:

```{r}

```


Plot the histogram of every variable in a facet plot with `ggplot2`. For that, you will have to convert the data frame into a **tidy data frame** (with the names of the variables in one column, and the corresponding values in another column):

```{r}

```



## Correlation

If you think any independent variable could be correlated with our dependent variable `y`, make a single scatter plot with `ggplot2` to take a deeper look:

```{r}

```



## Autocorrelation

Plot the **correlogram** of our target variable using function `ACF()` combined with `autoplot()`:

```{r}

```

Answer the following questions:

* Is there any autocorrelation?
* Is there any seasonality? Which periods do you detect?
* Is there any trend?

Try different time resolutions form the `index` in our tsibble, since the `lag` in the `ACF()` function depends on this.

```{r}

```

After analysing the trend and seasonality from the *correlogram*. Could you think on any other variable that could be interesting as predictor?

Add it in our `ev_ts` data frame:

```{r}

```



## Decomposition

Do you think that the decomposition will be relevant for our target variable?



Obtain the different decomposition components using functions `STL()` and `components()`, according to the different seasonality periods detected in last section. Then, plot the decomposition components using function `autoplot()`:

```{r}

```


Try applying now a logarithmic transformation to see if the decomposition changes:

```{r}

```

Does it have sense to apply a logarithmic transformation in our target variable?


